Titolo da decidere

### 2012-08-04 paper/paper.tex
analisi perf con esempio

\section{Example}\label{example}
L'esmepio scelto si ispira ad una procedura che trova applicazione all'interno di organizzazioni a carattere commerciale, si tratta del processo di vendita. In questa sede viene considerata solo una sua semplificazione rispetto a quanto avviene realmente. Per noi, il processo di vendita inizia con l'attività 'notify', in questa fase viene accolto un ordine da parte di un cliente. Dopo la ricezione dell'ordine, vengono condotte alcune attività in parallelo (visto che non presentano alcuna dipendenza tra loro) una di queste è l'attività 'FinancialCheck', in questa fase viene effettuata una valutazione dell'ordine dal punto di vista finanziario, possiamo ad esmepio pensare che vengono condotte delle indagini sulla solvibilità del cliente che ha emesso l'ordine. Parallelamente viene svolta l'attività 'ExternalWarehouse'...

@@

In general, there could be exist some transition in the net called {\itshape invisible}. This can happen if the transition models an internal choice that is not visible in the system, or it is used to implement a construct of a more abstract modelling language (BPMN or others). Usually, invisible transitions are considered to be lazy, i.e, when firing a visible transition $t$ corrisponding to an event of the trace, only then the invisibile transition enabling $t$ is fired. However, in this paper, we adopt an other way for handling these transitions and the firing is performed as soon as possible. This allows to carry out some performance measures as explained that ar not possible with the lazy metohode as explained in \cite.\\

@@

To clarify the conformance analysis technique we exploit the Petri net presented in section \ref{example} with reference to tow traces , $T$ and $T'$, presented respectively in figures \ref{ConfLog} and \ref{NonConfLog}. The log replay execution of the trace $T$, which is compliant with the Petri, terminates with a marking containing one token in the end place \{${p14 \rightarrow 1}$\}, and returns the sequence:
\begin{equation}
\begin{split}
R&=\{(Order\_start,1), (Order\_complete,2),(WarehouseCheck\_start,3), \\
& (FinancialCheck\_start,4),(WarehouseCheck\_complete,5),\\
& (FinancialCheck\_complete,6),(t10, 6), (Notify\_start,7), (Notify\_complete,8)\}
\end{split}
\end{equation}
The log replay execution of the trace $T'$ terminates with remaing tokens \{${p8 \rightarrow 1,p14 \rightarrow 1}$\} and a missing token \{${p10 \rightarrow 1}$\}, the sequence returned is:
\begin{equation}
\begin{split}
R'&=\{(Order\_start,1), (Order\_complete,2), (WarehouseCheck\_start), \\
& (FinancialCheck\_start,3), (WarehouseCheck\_start,4),(t7,5), (t10,6)\\
& (Notify\_start,6), (Notify\_complete,7)\}
\end{split}
\end{equation}

\begin{figure}[H]\label{ConfLog}
\centering
\includegraphics[width=400pt,height=40pt]
{./items/logConforme.pdf}
\caption{Conforme log example}
\end{figure}

\begin{figure}[H]\label{NonConfLog}
\centering
\includegraphics[width=400pt,height=40pt]
{./items/logNonConforme.pdf}
\caption{Non conforme log example}
\end{figure}

>>>>

\section{Example}\label{example}

\begin{figure}[t]\label{ConfLog}
\includegraphics[width=360pt]
{./items/Sales_PN.pdf}
\caption{Petri net}
\end{figure}

L'esmepio scelto si ispira ad una procedura che trova applicazione all'interno di organizzazioni a carattere commerciale, si tratta del processo di vendita. In questa sede viene considerata solo una sua semplificazione rispetto a quanto avviene realmente. Per noi, il processo di vendita inizia con l'attività 'notify', in questa fase viene accolto un ordine da parte di un cliente. Dopo la ricezione dell'ordine, vengono condotte alcune attività in parallelo (visto che non presentano alcuna dipendenza tra loro) una di queste è l'attività 'FinancialCheck', in questa fase viene effettuata una valutazione dell'ordine dal punto di vista finanziario, possiamo ad esmepio pensare che vengono condotte delle indagini sulla solvibilità del cliente che ha emesso l'ordine. Parallelamente viene svolta l'attività 'ExternalWarehouse'...

@@

In general, there could be exist some transition in the net called {\itshape invisible}. This can happen if the transition models an internal choice that is not visible in the system, or it is used to implement a construct of a more abstract modelling language (BPMN or others). Usually, invisible transitions are considered to be lazy, i.e, when firing a visible transition $t$ corrisponding to an event of the trace, only then the invisibile transition enabling $t$ is fired. However, in this paper, we adopt an other way for handling these transitions and the firing is performed as soon as possible. This allows to carry out some performance measures as explained that are not possible with the lazy metohode as explained in \cite.(riferimento al paper Applying Process Analysis to the italian eGiv Entreprise Architecture)\\

@@

To clarify the conformance analysis technique we exploit the Petri net presented in section \ref{example} with reference to tow traces , $T$ and $T'$, presented respectively in figures \ref{ConfLog} and \ref{NonConfLog}. The log replay execution of the trace $T$, which is compliant with the Petri net, terminates with a marking containing one token in the end place \{${p14 \rightarrow 1}$\}, and returns the sequence:
\begin{equation}
\begin{split}
R&=\{(Order\_start,1), (Order\_complete,2),(WarehouseCheck\_start,3), \\
& (FinancialCheck\_start,4),(WarehouseCheck\_complete,5),\\
& (FinancialCheck\_complete,6),(t10, 6), (Notify\_start,7), (Notify\_complete,8)\}
\end{split}
\end{equation}
The log replay execution of the trace $T'$ terminates with remaing tokens \{${p8 \rightarrow 1,p14 \rightarrow 1}$\} and a missing token \{${p10 \rightarrow 1}$\}. The missing token is created artificially and this fact records a wrong execution of the event $Notify\_start$. In fact, notice that in the trace $T'$ this event is executed before the termination of the activity ``FinancialCheck'', and this is interpreted as a non conformance to the process model. The sequence returned by log replay is:
\begin{equation}
\begin{split}
R'&=\{(Order\_start,1), (Order\_complete,2), (WarehouseCheck\_start), \\
& (FinancialCheck\_start,3), (WarehouseCheck\_start,4),(t7,5), (t10,6)\\
& (Notify\_start,6), (Notify\_complete,7)\}
\end{split}
\end{equation}

\begin{figure}[t]\label{ConfLog}
\centering
\includegraphics[width=400pt,height=40pt]
{./items/logConforme.pdf}
\caption{Conforme log example}
\end{figure}

\begin{figure}[t]\label{NonConfLog}
\centering
\includegraphics[width=400pt,height=40pt]
{./items/logNonConforme.pdf}
\caption{Non conforme log example}
\end{figure}

Since the logs contain timestamps, the log replay can be used to compute performance measures of the process. The idea is to calculate the time interval between production and consumption of tokens in each place. This tecniques can be applied only to traces that do not present missing token during the replay, because such tokens cannot have time information. During the log replay the following metrics can be computed for catch trace and each place:
\begin{itemize}
\item sejour time $(tsj)$ : the time interval between arrival and depature of tokens;
\item synchronization time $(tsc)$: the time interval between arrival of a token in the place and enabling of a transition in the post-set of the place;
\item waiting time $(tw)$:  the time interval between enabling of a transition in the post-set of the place and token departure (thus $tsj=tsc+tw $).
\end{itemize}

To clarify the metrics evaluated by this tecnique, as done with conformance analysis, we exploit the Petri net  presented in section \ref{example} and the trace T presented in figure \ref{ConfLog}. The replay starts with firing the transition $Order\_start$ at time $0s$, thus a token arrives at $p2$ at time $0s$. After $45 min$ the token in $p2$ is consumed and the transition $Order\_complete$ is fired, so $tw(p2)=tsj(p2)=45min$, and a token arrives at $p3$ and $p4$. At the time $1h:5min:25s$ the transition $WarhouseCheck\_start$ is fired, thus $tw(p4)=20min+25sec$ and a token is produced at $p6$. At the time $1h:10min:20s$ the transition $FinancialCheck\_start$ is fired and a token is produced at $p5$, so $tw(p3)=25min+20s$. After $2h:30min:20s$ from the replay start the transition $Financial\_complete$ is fired, hence we have $tw(p5)=1h+20min$ and a token is produced at $p11$. At time $3h:10min:20s$ the firing of the transition $WarehouseCheck\_complete$ is executed, so $tw(p6)=2h+4min+55s$ and token is produced at $p7$ and $p8$. At the same time the invisible transition $t10$ is fired and a token is produced at $p10$. At this point the transition $t10$ is enabled, so firing it is now possible and a token is produced at $p12$. Notice that $tsc(p11)=40min, tsc(p10)=0s$. At the time $3h:25min:20s$ the transition $Notify\_start$ is fired and a token is produced in $p13$, after $4min+4sec$ the trasition $Notify\_complete$ is fired and $tw(p13) = 4min+40s$.
